---
title: "Feasibility analysis of Yahoo acquistion of Tumblr"
date: "January 15, 2019"
output:
  slidy_presentation: default
  beamer_presentation: default
Author: SingRs (ASHRAF Shahan, BRITO DA FONSECA Fabio, HARRIS Alana, MILENKOV Dimitar,
  OBOLONSKYI Misha)
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(forecast)
library(fpp)
library(xts)
library(scales)
```

##Deliverables
- What was the average monthly growth rate in the number of people worldwide accessing Tumblr site since its inception (over the last 37 months)?
- How does it compare to the average monthly growth rate over the past 12 months?
- What are the valuations of Tumblr given these two growth rates?
- Use the data provided to forecast the number of people worldwide accessing Tumblr's site for the next 115 months (June 2013 - Dec 2022).
- Given your forecast, revise the valuation of Tumblr (case Exhibit 5, valuation spreadsheet).

##Table of Context

1.Reading data  
2.Average growth rates  
3.Valuation comparisons  
4.Decomposing data  
5.Forecasting  
6.TBATS  
7.Arima  


##Reading data  
```{r, comment="", echo = TRUE}
tumblr_visitors <- read.csv(file.choose(), header=TRUE, sep=",")
tumblr_visitors_ts <- ts(tumblr_visitors$People, start = c(2010, 04), frequency = 12)
```

##Explonatory analysis (1/2)
```{r, comment="", echo = TRUE}
summary(tumblr_visitors_ts)
head(tumblr_visitors_ts[1:5])
tail(tumblr_visitors_ts[34:38])
```

##Explonatory analysis (2/2)
```{r, comment="", echo = TRUE}
plot(as.xts(tumblr_visitors_ts/1000000), ylab="Millions of visitors on Tumblr website",
     major.format = "%Y-%m", main="Tumblr monthly users, mln")
```

##Average growth rates over the past 37 vs 12 months
- Growth rate for the past 37 months
```{r, comment="", echo = TRUE}
growth_rates_cagr_37 <- (tumblr_visitors_ts[38]/tumblr_visitors_ts[1])^(1/38)-1
```
- Growth rate for the past 12 months
```{r, comment="", echo = TRUE}
growth_rates_cagr_12 <- (tumblr_visitors_ts[38]/tumblr_visitors_ts[38-12])^(1/12)-1
```
- creating matrix for comparison
```{r, comment="", echo = TRUE}
val_sensitivy <- data.frame("Growth rate" = c(growth_rates_cagr_12, growth_rates_cagr_37),
                            "Valuation"=as.numeric(c("1.23","54.8")))
```

##Tumblr valuation dependeding on people growth rate
```{r, comment="", echo = TRUE, fig.height=6}
val <- barplot(val_sensitivy$Valuation, names.arg = percent(val_sensitivy$Growth.rate),
        xlab = "Growth rates", ylab = "Valuation of Tumblr, bln USD", ylim = c(0,60))
text(x=val, y=val_sensitivy$Valuation, labels=as.character(val_sensitivy$Valuation), 
     pos = 3, cex = 1)
```

##Decomposition of Tumblr visitors - Multiplicative
```{r, comment="", echo = TRUE}
fit <- decompose(tumblr_visitors_ts, type="multiplicative") 
plot(fit)
```

##Decomposition of Tumblr visitors - Additive
```{r, comment="", echo = TRUE}
fit <- decompose(tumblr_visitors_ts, type="additive") 
plot(fit)
```

##Decomposition of Tumblr visitors - STL (Season and trend using Loess)
```{r, comment="", echo = TRUE}
fit <- stl(tumblr_visitors_ts, t.window=12, s.window="periodic") 
plot(fit)
```


##Modelling
#### Create exponential smoothing models: additive vs multiplicative noise (first A vs M), additive vs multiplicative trend (second A vs M) and no seasonality vs automatic detection (third N vs Z) trend and no seasonlity (AAN), multiplicative (MMN)
```{r, comment="", echo = TRUE}
tumblr_visitors_AAN <- ets(tumblr_visitors_ts, model="AAN")
tumblr_visitors_AAZ <- ets(tumblr_visitors_ts, model="AAZ", damped=FALSE)
tumblr_visitors_MMN <- ets(tumblr_visitors_ts, model="MMN", damped=FALSE)
tumblr_visitors_MMZ <- ets(tumblr_visitors_ts, model="MMZ", damped=FALSE)
```

#### Create their prediction "cones" for 115 months into the future with quintile confidence intervals
```{r, comment="", echo = TRUE}
tumblr_visitors_AAN_pred <- forecast(tumblr_visitors_AAN, h=115, level=c(0.8, 0.95))
tumblr_visitors_AAZ_pred <- forecast(tumblr_visitors_AAZ, h=115, level=c(0.8, 0.95))
tumblr_visitors_MMN_pred <- forecast(tumblr_visitors_MMN, h=115, level=c(0.8, 0.95))
tumblr_visitors_MMZ_pred <- forecast(tumblr_visitors_MMZ, h=115, level=c(0.8, 0.95))
```

##Compare the prediction "cones" visually
```{r, comment="", echo = TRUE}
par(mfrow=c(1,4))
plot(tumblr_visitors_AAN_pred, xlab="Year")
plot(tumblr_visitors_AAZ_pred, xlab="Year") 
plot(tumblr_visitors_MMN_pred, xlab="Year")
plot(tumblr_visitors_MMZ_pred, xlab="Year")
```

##Trigonometric box-cox autoregressive trend seasonality (TBATS) model
```{r, comment="", echo = TRUE}
tumblr_visitors_tbats <- tbats(tumblr_visitors_ts)
tumblr_visitors_tbats_pred <-forecast(tumblr_visitors_tbats, h=115, level=c(0.8, 0.95))
par(mfrow=c(1,1))
plot(tumblr_visitors_tbats_pred, xlab="Year")
```


##Arima
```{r, comment="", echo = TRUE}
fit_non_seasonal <- auto.arima(tumblr_visitors_ts,seasonal=FALSE)
fit_seasonal <-auto.arima(tumblr_visitors_ts,seasonal=TRUE)
#automatically fits the ARIMA model (auto-regressive integrated moving average)
par(mfrow=c(1,2))
Acf(residuals(fit_non_seasonal))
acf(residuals(fit_seasonal))
```

##Arima plots
```{r, comment="", echo = TRUE}
par(mfrow=c(1,2))
plot(forecast(fit_non_seasonal,115))
plot(forecast(fit_seasonal,115))
```

##Best model

- bullet 1
- bullet 2
- bullet 3

##Comparing valuation with forecasted growth rates